My Data Literacy Project\\ (Replace this with your Project Title)

### 2024-01-16 doc/DataLiteracyStudentProject/paper.tex
updated paper

\usepackage{tikz}
% Corporate Design of the University of Tübingen
% Primary Colors
\definecolor{TUred}{RGB}{165,30,55}
\definecolor{TUgold}{RGB}{180,160,105}
\definecolor{TUdark}{RGB}{50,65,75}
\definecolor{TUgray}{RGB}{175,179,183}

@@

%Put your abstract here. Abstracts typically start with a sentence motivating why the subject is interesting. Then mention the data, methodology or methods you are working with, and describe results. 
\begin{abstract}
As often mentioned in the news and in the daily life, people suggest that a higher 
\end{abstract}
%Motivate the problem, situation or topic you decided to work on. 
% Describe why it matters (is it of societal, economic, scientific value?). 
% Outline the rest of the paper (use references, e.g.~to \Cref{sec:methods}: 
% What kind of data you are working with, how you analyse it, and what kind of conclusion you reached. 
% The point of the introduction is to make the reader want to read the rest of the paper.

\section{Introduction}\label{sec:intro}
As often mentioned in the news and in the daily life, people suggest that a higher expected salary influences the job aspirations of the youth. 
We are curious about the underlying truth of thgis stetement. We have three Datasets, one Datset shows the Expected Salaries for various industrial sectors,
the second one shows the amount of students in the STEM sector in Tübingen from the Wintersemester 2005/2006 to the Wintersemester 2023/2024. 
The third dataset contains information about the amount of students, who finish high school in the certain year.
We did a feature analysis of the three datasets to get a better understanding of the data but the main focus of our analysis is on predicting 
the amount of students in the STEM sector for the next years. We used different approaches like a Random Forest approach with SARIMA
and a Multivariate Autoregression Model to investigate which approach works best for our purpose.

% In this section, describe \emph{what you did}. Roughly speaking, 
% explain what data you worked with, how or from where it was collected, it's structure and size.
%  Explain your analysis, and any specific choices you made in it. Depending on the nature of your project,
% you may focus more or less on certain aspects. If you collected data yourself, explain the collection process in detail.
% If you downloaded data from the net, show an exploratory analysis that builds
% intuition for the data, and shows that you know the data well. 
% If you are doing a custom analysis, explain how it works and why it is the right choice.
% If you are using a standard tool, it may still help to briefly outline it. Cite relevant works. 
% You can use the \verb|\citep| and \verb|\citet| commands for this purpose \citep{mackay2003information}.
\section{Data and Methods}\label{sec:methods}
We used three different pre collected datasets, our University data was collected by the Statistics Department from the University of Tuebingen. 
They used the immatriculation data from the students to create this dataset, which leads to
a quite accurate data which is necessary for the usage in our models.
\textcolor{red}{// HERE WE NEED MORE INFORMATION ABOUT THE DATA FROM THE OTHER DATASETS//}\\
The first step was to do some basic feature engineering on the datasets. We used various plots like line graphs, bar charts and pie charts
to get a better understanding of the data and detect trends. 
After we finished the basic feauture engineering we started the main analysis of the Vector Autoregressive Model (VAR). 
To get the most of this model, we wrote a parser, which parses text data generated from a PDF. This parser basically uses regular expressions to extract the amount of students
per years divided into the semesters and created a new dataset with the usage of pandas.\\
We've chosen the vector autoregressive model, because we have multivariate data and want to get a deeper understanding, if one of the datasets affects the others, 
so that we are able to predict one time series with another. A vector autoregressive model doesn't predict a single time series analyses, 
but multiple. Each time series is a linear combination which consists of n lags and an error term. A lag is the data from the past which we use to estimate
a new data point. We define the amount of lags in the hyperparameter p which defines the maxamount of lags which are used in the linear combination.
This linear combination also uses an error term, because we can't guarantee that our model predicts the correct output.\\
\textcolor{red}{One crucial step in the usage of those AR models, is that we have to ensure that our data is stationary. 
Data is stationary if those three attributes are fulfilled
\begin{enumerate}
    \item The mean $\mu$ is constant
    \item The standard deviation is constant
    \item There is no seasonality in the data
\end{enumerate}
// CHECK IF OUR DATA FULFILLS THOSE THREE ATTRIBUTES ///\\
After we ensured this, we did a PCAF to find statistically significant lags which we could use to find an optimal hyperparameter p of our vector autoregressive model.
A PCAF only represents the direct influence of the past datapoint $S_{n-i}$ for $i \in \{0,1,...,n\}$  to $S_n$. This gives us a graphical visualization with an error band of non statistical significant
lags in // GIVE A RANGE //.\\
After we used this PCAF we tried the different amount of lags, which looked like they have a significant input on the current datapoint $S_n$.
With the optimal lag we did a Granger Causality Test to get an understanding if one timeseries really influences another. The Granger Causality Test, does a t-Test and F-Test
checks if we can predict $S_i$ with $i \in \{2,3\}$ the time series $S_j$ $i\neq j$. The Granger Causality Test internally injects lags from the time series $S_J$ 
into the linear combination of $S_i$, then it does a t-Test and an F-test. As soon as one of those terms is verified by both tests,
then the test result is true.
}

>>>>

\usepackage{tikz}
\usepackage{enumitem}
% Corporate Design of the University of Tübingen
% Primary Colors
\definecolor{TUred}{RGB}{165,30,55}
\definecolor{TUgold}{RGB}{180,160,105}
\definecolor{TUdark}{RGB}{50,65,75}
\definecolor{TUgray}{RGB}{175,179,183}

@@

%Put your abstract here. Abstracts typically start with a sentence motivating why the subject is interesting. Then mention the data, methodology or methods you are working with, and describe results. 
\begin{abstract}
As often mentioned in the news and in the daily life, people suggest that a higher expected salary is a strong pull factor for students 
choosing their subject. We were curious if this is particullarily true for the scientific based subjects, because a majority of them have a 
pretty high expected salary. We used the students data from the \href{https://uni-tuebingen.de/einrichtungen/verwaltung/iv-studierende/studierendenabteilung/statistiken/}{Statistics University of Tübingen}, 
the high-school graduates data from the \href{https://www.statistik-bw.de/BildungKultur/SchulenAllgem/LRt0302.jsp}{Statistisches Landesamt Baden-Württemberg}
and also the salary data from 2005 to 2023 from the \href{https://www-genesis.destatis.de/genesis//online?operation=table&code=62321-0001&bypass=true&levelindex=0&levelid=1702307320529#abreadcrumb}{Deutsches Statistisches Bundesamt}.
We used a vector autoregressive model, to build a time series forecast and analysed the outcomes to find an optimal configuration for this type of model in our context.
\textcolor{red}{result.}
%Motivate the problem, situation or topic you decided to work on. 
% Describe why it matters (is it of societal, economic, scientific value?). 
% Outline the rest of the paper (use references, e.g.~to \Cref{sec:methods}: 
% What kind of data you are working with, how you analyse it, and what kind of conclusion you reached. 
% The point of the introduction is to make the reader want to read the rest of the paper.

\section{Introduction}\label{sec:intro}
In the public eye, there is a strong correlation between the expected salary and the subject enrollment of students. We are curious if this statement is true or just a biased
perception of the society. Our main focus is on the University of Tübingen and mainly on the STEM (scientific engineering technical and mathematical) subjects, because those degrees 
are likely to have a pretty high expected salary after graduating. Furthermore we needed data of high-school graduates in Baden-Würrtemberg, to find a possible corellation between
the amount of graduates and the new enrollments in the stem subjects. We just focused on Baden-Württemberg, this introduces a bias to our prediction, but we assumed that the 
majority of students which are going to study in Tübingen come from Baden-Württemberg. The third datatset we used is about the salaries from 2005 to 2023 in the main 
industrial sectors in Germany. The main focus of this paper is on a analysis of a vetor autoregressive model, to predict the upcoming numbers of students in the next semesters. We
used various techniques to clean our data like removing the inflation or making our time series stationary to ensure that the predictions don't detect patterns which are inferred by 
some noise in the data. Nevertheless we also used basic data science techniques to get a better understanding of the dirstributions and trends over the past years.

% In this section, describe \emph{what you did}. Roughly speaking, 
% explain what data you worked with, how or from where it was collected, it's structure and size.
%  Explain your analysis, and any specific choices you made in it. Depending on the nature of your project,
% you may focus more or less on certain aspects. If you collected data yourself, explain the collection process in detail.
% If you downloaded data from the net, show an exploratory analysis that builds
% intuition for the data, and shows that you know the data well. 
% If you are doing a custom analysis, explain how it works and why it is the right choice.
% If you are using a standard tool, it may still help to briefly outline it. Cite relevant works. 
% You can use the \verb|\citep| and \verb|\citet| commands for this purpose \citep{mackay2003information}.
\section{Data and Methods}\label{sec:methods}
In our research we used three datasets, which are divided as follows: The first dataset is from the \href{https://uni-tuebingen.de/einrichtungen/verwaltung/iv-studierende/studierendenabteilung/statistiken/}{Statistics Bureau of theUniversity of Tübingen}., which contains the
total amount of students from the winter semester 2005/2006 to the winter semester 2023/2024. The data was collected from the enrollments of students, which registered 
for a study in Tübingen.
The second dataset is from the \href{https://www.statistik-bw.de/BildungKultur/SchulenAllgem/LRt0302.jsp}{Statistisches Landesamt Baden-Württemberg} which contains data about the high-school graduates from 1970 to 2023.
Our third and last dataset is from the \href{https://www-genesis.destatis.de/genesis//online?operation=table&code=62321-0001&bypass=true&levelindex=0&levelid=1702307320529#abreadcrumb}{Deutsches Statistisches Bundesamt} and contains the salaries of the large industrial sectors in Germany from 2005 to 2021. 
One main problem with the mapping between university subjects and actual jobs is that there exists no perfect one-to-one mapping. 
We also removed the inflation from the salaries to get the real wage, because the inflation could 
cause a false pattern detection in our autoregressive model. 
To reduce the amount of bias introduced in this step, we also used the Inflation data from the Deutsches Statistisches Bundesamt over the years.\\

\textcolor{red}{To get the most of this model, we wrote a parser, which parses text data generated from a PDF. This parser basically uses regular expressions to extract the amount of students
per years divided into the semesters and created a new dataset with the usage of pandas.\\}

Our first steps involved some feauture engineering on the three different datasets to get a better understanding of the data and also detect patterns in the data like trends and distributions over time.
After we finished the basic feauture engineering step we started the main analysis of the Vector Autoregressive Model (VAR). 
We've chosen the vector autoregressive model, because we have multivariate data and we assumes that there is a correlation between the given time series.
A VAR model don't predict a single time series but one for each input series. Those modles have one hyperparameter \textcolor{red}{p} which define the amount 
of previous data points which should be represented in the output series. Each prediction is a linear combination with the form  $S_n = C_{1} \cdot S_{n-1}+ C_{2} \dots+ C_i \cdot  S_{n-i} + \epsilon$ for
$i \in \{1,...,|lags|\}$. In our case we have three incoming time series which also leads to three outgoing time series. Our outgoing prediction for the data point n based on lags 
is represented as
\begin{equation*}
    \resizebox{\linewidth}{!}{%
        $\begin{bmatrix}
            S_{1,n}\\
            S_{2,n}\\
            S_{3,n}
        \end{bmatrix} = 
        \begin{bmatrix}
            C_{1,1} & C_{1,2} & \dots &C_{1,n-1}\\
            C_{2,1} & C_{2,2} &\dots & C_{2,n-1}\\
            C_{3,1} & C_{3,2} &\dots &C_{3,n-1}\\
        \end{bmatrix} + 
        \begin{bmatrix}
            S_{1,1} & S_{1,2} & \dots & S_{1,n-1}\\
            S_{2,1} & S_{2,2} & \dots & S_{2,n-1}\\
            S_{3,1} & S_{3,2} & \dots & S_{3,n-1}\\
        \end{bmatrix}+
        \begin{bmatrix}
            \epsilon_{1}\\
            \epsilon_{2}\\
            \epsilon_{3}
        \end{bmatrix}$%
    }
    \end{equation*}
Each time series $S_n$ $\forall i \in \{1,2,3\}$ describes the n-th data point for the input time series i. The term $\epsilon$ is added to approximate an error, which is 
induced by the model.\\
We need to assure that our model is stationary, which means that the the stochastic process doesn't change over time. \textcolor{red}{WHY?}.
Our data is stationary if we are able to assure this for all our time series data inputs:
\begin{enumerate}[label=\Roman*]
    \item The mean $\mu$ is constant
    \item The standard deviation is constant
    \item There is no seasonality in the data
\end{enumerate}

After we ensured, that our data is stationary, we used a PACF plot to find lags which could be statistically significant. 
The main reason why we choose PACF in contrast to ACF is that we only want to look at direct influences from $S_{n-1}$ to $S_{n}$ and not from $S_{n-3}$ to $S_n$.
\begin{figure}[H]
    \centering
    \includegraphics[width=\linewidth,height=\textheight, keepaspectratio]{fig/pacf_plot.pdf}
    \caption{PACF Plot for our VAR Model}
    \label{fig:enter-label}
\end{figure}
\textcolor{red}{// CONCLUSION FOR PACF // }\\

After we found the optimal lag configuration of our model, we used a Granger Causality Test, to statistically prove that the salary time series really influences the student decision.
The main idea of Granger Causality is that we are able to predict the data point $S_{i,n}$ directly with one or multiple other time serieses. The Granger Causality Test uses exactly this
idea, by inserting data points $S*:=S_{j,m}$ for $j\in i:j\neq i$  and $m<n$. After we inserted a couple of terms we do a t-Test and an F-Test on this new $S*$ and as soon as both tests
return statistical significance for the same S* we can assure that the three time series are Granger Causal.
